Unsupervised embeddings of organizations capture the latent structure of scientific mobility

### 2020-06-16 paper/current/main.tex

Human mobility and migration drives countless social phenomenon including the growth and evolution of cities~\autocite{bettencourt2007cities, bettencourt2013cities, wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity, hufnagel2004epidemics, eubank2004outbreaks}, economies~\autocite{kaluza2010cargo, kerr2011immigration, taylor2016refugees, pappalardo2015mobility}, innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation, liu2010spillovers},  and scientific research~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity, simini2012universal}, describes mobility using the \textit{distance} between places.
However, amid a globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization}, geographic distance has less relevance for understanding mobility, whereas other abstract notions of distance~\autocite{boschma2005proximity} pose practical and theoretical challenges.
Using the case of scientific mobility, we leveragethe neural embedding model~\autocite{mikolov2013word2vec} to demonstrate a data-driven approach for learning a vector-space embedding that encodes a functional distance~\autocite{brown1970migration, brown1970functional, kim2018functional} between places and captures the latent and multi-faceted structure of human mobility.
We train embeddings of scientific organizations using mobility trajectories derived from the publications of disambiguated authors from the Web of Science database between 2008 and 2019.
Distance between vectors in the learned embedding space explains and predicts real mobility between organizations better than  geographic distance. 
We argue that this explanatory power emerges from an underlying connection between the neural embedding model and the gravity model.
In addition to geography, the embedding also captures latent factors related to culture, language, and prestige of organizations at multiple levels of granularity.
With this approach, we can learn scalable, dense, and meaningful vector representations of scientific mobility directly from the data using a methodology that can be easily applied to all kinds of mobility data.
\end{abstract}

@@

Here, we propose a neural embedding model that uses real-world mobility trajectories to encode the complex structure of mobility into a dense, continuous, and meaningful vector-space representation, from which a \textit{functional distance} between places can be derived.
This approach, origionally designed to create embeddings of words~\autocite{mikolov2013word2vec} has been used to reveal insights in texts~\autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}.
In addition to providing a data-driven functional distance, this approach is computationally efficient, scalable for large datasets, robust to noise, and allows for a range of unique analyses that leverage their semantic properties using simple vector arithmetic.

>>>>

Human mobility and migration drives major social phenomenon such as the growth and evolution of cities~\autocite{bettencourt2007cities, bettencourt2013cities, wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity, hufnagel2004epidemics, eubank2004outbreaks}, economies~\autocite{kaluza2010cargo, kerr2011immigration, taylor2016refugees, pappalardo2015mobility}, innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation, liu2010spillovers},  and scientific research~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity, simini2012universal}, describes mobility using the \textit{distance} between places.
However, amid a globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization}, geographic distance has less relevance for understanding mobility, whereas other abstract notions of distance~\autocite{boschma2005proximity} pose practical and theoretical challenges.
Using the case of scientific mobility, we leverage the neural embedding model~\autocite{mikolov2013word2vec} to demonstrate a data-driven approach for learning a vector-space embedding that encodes a functional distance~\autocite{brown1970migration, brown1970functional, kim2018functional} between places and captures the latent and multi-faceted structure of human mobility.
We train embeddings of scientific organizations using mobility trajectories derived from the publications of disambiguated authors from the Web of Science database between 2008 and 2019.
Distance between vectors in the learned embedding space explains and predicts real mobility between organizations better than  geographic distance, emerging from an underlying connection between the neural embedding model and the gravity model.
Exploiting their unique properties, we interpret the embedding space and observe how it captures latent factors related to culture, language, and prestige of organizations at multiple levels of granularity.
With this approach, we can learn scalable, dense, and meaningful vector representations of scientific mobility directly from the data using a methodology that can be easily applied to all kinds of mobility data.
\end{abstract}

@@

Here, we propose a neural embedding model that uses real-world mobility trajectories to encode the complex structure of mobility into a dense, continuous, and meaningful vector-space representation, from which a \textit{functional distance} between places can be derived.
This approach, origionally designed to create embeddings of words~\autocite{mikolov2013word2vec} has been used to reveal insights in texts~\autocite{linzhuo2020hyperbolic, tshitoyan2019mat2vec, garg2018gender, kozlowski2018geometry, hamilton2016diachronic}, documents~\autocite{le2014doc2vec, nakandala2016twitch}, and networks~\autocite{perozzi2014deepwalk, grover2016node2vec}.
In addition to providing a data-driven functional distance, this approach is computationally efficient, scalable for large datasets, robust to noise, and are readily interpretable using an array of unique techniques based on simple vector arithmetic.
